---
title: "Statistical consulting: Disney World case"
author: "Raïsa Carmen"
date: "`r Sys.Date()`"
tables: true
graphics: yes
output:
  beamer_presentation:
    theme: "Hannover"
    colortheme: "seahorse"
    fonttheme: "professionalfonts"
    toc: TRUE
    slide_level: 2
    keep_tex: true
    includes:
header-includes: |
  \usepackage{etoolbox}

  \defbeamertemplate{background canvas}{mydefault}{%
    \includegraphics[scale = 0.1, trim=-106.7cm 0 0 -79.6cm]{disney-channel-the-walt-disney-company-toon-disney-clip-art-grey-wallpaper.jpg}
  }
  \defbeamertemplate{background canvas}{standout}{%
    
  }

  \BeforeBeginEnvironment{frame}{%
    \setbeamertemplate{background canvas}[mydefault]%
  }

  \makeatletter
  \define@key{beamerframe}{standout}[true]{%
    \setbeamertemplate{background canvas}[standout]%
  }
  \makeatother
---

```{r setup, include=FALSE}
# choose a beamer theme, font and color
# https://deic.uab.cat/~iblanes/beamer_gallery/
# beamer : markdown
# https://bookdown.org/yihui/rmarkdown/beamer-presentation.html
knitr::opts_chunk$set(echo = FALSE)
```

```{r loaddata, include=FALSE, message=FALSE, warning=FALSE, appendix=TRUE}
#---------------------------load all packages----------------------------------#
library(tidyverse) # data wrangling and visualisation
library(rprojroot) # relative file paths
library(kableExtra)# tables in markdown
library(git2rdata) # clean, git-friendly datasets
library(patchwork) # combining several plots into one
library(brms)      # fitting Bayesian models
library(bayesplot) # visualise MCMC and posterior predictive checks
library(tidybayes) # wrangling and visualising Bayesian models

# Conflicts between packages
conflicted::conflicts_prefer(dplyr::filter)
conflicted::conflicts_prefer(dplyr::lag)
conflicted::conflicts_prefer(brms::ar)
conflicted::conflicts_prefer(brms::dstudent_t)
conflicted::conflicts_prefer(brms::pstudent_t)
conflicted::conflicts_prefer(brms::qstudent_t)
conflicted::conflicts_prefer(brms::rstudent_t)
conflicted::conflicts_prefer(dplyr::pull)
conflicted::conflicts_prefer(dplyr::group_rows)
conflicted::conflict_prefer("rhat", "brms")

knitr::opts_chunk$set(echo = FALSE, message = FALSE)
knitr::opts_knit$set(
  root.dir = find_root(criterion = has_file("stat_cons_disney.Rproj")))

#----------------------------Load all necessary data---------------------------#
waiting_times <- read_vc(file = "waiting_times",
                         root = find_root_file("data",
                                               criterion =
                                                 has_file(
                                                   "stat_cons_disney.Rproj"))) #SPOSTMIN:  Value -999 indicates the attraction was closed
file_to_ride <- read_vc(file = "file_to_ride",
                        root = find_root_file("data",
                                              criterion =
                                                has_file(
                                                  "stat_cons_disney.Rproj")))
entities_extra <- read_csv(find_root_file("data", "entities_extra.csv",
                                              criterion =
                                                has_file(
                                                  "stat_cons_disney.Rproj")))
metadata <- read_csv(find_root_file("data", "metadata.csv",
                                              criterion =
                                                has_file(
                                                  "stat_cons_disney.Rproj")))
file_to_ride <- file_to_ride %>%
  dplyr::filter(!is.na(code)) %>%
  arrange(file, opened_on) %>%
  group_by(file, short_name) %>%
  slice_tail() %>%#keep only the last occurance
  ungroup()
#keep only waiting times with corresponding posted waiting time:
waiting_times <- waiting_times %>%
  dplyr::filter(!is.na(posted_at_start_wait) &  posted_at_start_wait != -999 &
                  SACTMIN >= 0 & SACTMIN <= 360) %>%
  mutate(diff = posted_at_start_wait - SACTMIN)
set.seed(2024)
sample <- waiting_times %>%
  slice_sample(n = 10000)
```

# Introduction

- Waiting times for different attractions in Disney World
  - posted waiting times
  - actual waiting times
- Metadata
  - on the attractions
  - on each of the days



## Research questions

- How much does the posted waiting time differ from the actual waiting time?
- What influences the accuracy of the posted waiting times?
- How does the accuracy of the posted waiting times differ between attractions? 


## Assumptions

```{r echo= FALSE, include=TRUE}
datafolder <- find_root_file("data",
                             criterion = has_file("stat_cons_disney.Rproj"))
data_wait_7dwarfs <- read.csv(file = paste0(datafolder,
                                                "/waiting times/",
                                                "7_dwarfs_train.csv")) %>%
  dplyr::select(datetime, SACTMIN, SPOSTMIN)
data_wait_7dwarfs[255:265,] %>%
  kable(format = "latex", row.names = FALSE) %>%
  kableExtra::kable_styling(font_size = 9)

waiting_times %>%
  dplyr::select(datetime, SACTMIN, posted_at_start_wait) -> t
t[2,] %>%
  kable(format = "latex", row.names = FALSE) %>%
  kableExtra::kable_styling(font_size = 9)
```

# Data exploration

## Waiting times 


```{r}
p1 <- sample %>%
  ggplot() +
  geom_line(data = data.frame(x = c(0, max(sample$SACTMIN)),
                              y = c(0,max(sample$posted_at_start_wait))),
            aes(x = x, y = y)) +
  geom_point(aes(x = SACTMIN, y = posted_at_start_wait, color = diff),
             alpha = 0.5) +
  scale_color_gradientn("Difference",
                        colours =
                          c("yellow", "orange", "red", "darkred")) +
  xlab("Actual waiting time (min)") +
  ylab("Posted waiting time (min)") +
  guides(colour = "none") +
  theme_bw()

bin_width <- 10
nbins = ceiling((max(sample$diff) - min(sample$diff))/bin_width)
colfunc <- colorRampPalette(c("yellow", "orange", "red", "darkred"))

p2 <- sample %>%
  ggplot() +
  geom_histogram(aes(x = diff, y = after_stat(count / sum(count))),
                 binwidth = bin_width,
                 fill = colfunc(nbins),
                 alpha = 0.8) +
  geom_vline(aes(xintercept = 0), color = "black") +
  scale_y_continuous(labels = scales::percent) +
  ylab("Percentage of observations") +
  xlab("Posted - actual waiting time (min)") +
  theme_bw()

p4 <- waiting_times %>%
  mutate(
    start_wait = datetime - as.difftime(SACTMIN, units = "mins"),
    hour = hour(start_wait)) %>%
  group_by(hour) %>%
  summarize(mean = mean(diff),
            st_dev = sd(diff),
            nb = n(),
            lower = mean - 1.95 * st_dev/sqrt(nb),
            upper = mean + 1.95 * st_dev/sqrt(nb)) %>%
  ungroup() %>%
  filter(hour >= 6 & hour <= 23) %>%
  ggplot(aes(x = hour, y = mean)) +
  geom_ribbon(aes(ymin = lower, ymax = upper),
              alpha = 0.5) +
  geom_line() +
  theme_bw() +
  scale_x_continuous(breaks = seq(6, 23, 3)) +
  ylab("Posted - actual waiting time (min)") +
  xlab("Time of day (hour)")
p3 <- waiting_times %>%
  mutate(
    start_wait = datetime - as.difftime(SACTMIN, units = "mins"),
    hour = hour(start_wait)) %>%
  group_by(hour) %>%
  summarize(mean = mean(SACTMIN),
            st_dev = sd(SACTMIN),
            nb = n(),
            lower = mean - 1.95 * st_dev/sqrt(nb),
            upper = mean + 1.95 * st_dev/sqrt(nb)) %>%
  ungroup() %>%
  filter(hour >= 6 & hour <= 23) %>%
  ggplot(aes(x = hour, y = mean)) +
  geom_ribbon(aes(ymin = lower, ymax = upper),
              alpha = 0.5) +
  geom_line() +
  theme_bw() +
  scale_x_continuous(breaks = seq(6, 23, 3)) +
  ylab("Actual waiting time (min)") +
  xlab("Time of day (hour)")


p1 + p2 + p3 + p4

```


# Methodology and results

## Model

Linear, Bayesian model to predict the difference between the posted and actual waiting time.

```{r covariates}
waiting_times_reg <- waiting_times %>%
  left_join(metadata %>%
              dplyr::select(DATE, DAYOFWEEK, YEAR, HOLIDAYPX, WDW_TICKET_SEASON,
                            WDWMEANTEMP, inSession, MKPRDDAY,
                            CapacityLostWGT_MK, CapacityLostWGT_AK,
                            CapacityLostWGT_EP, CapacityLostWGT_HS) %>%
              mutate(WDW_TICKET_SEASON = ifelse(is.na(WDW_TICKET_SEASON),
                                                "regular", WDW_TICKET_SEASON)),#assume NA = "regular"-> should adjust this in metadata before join (NA here then means that date wasn't available in the metadata file)
            by = join_by(date == DATE)) %>%
  left_join(file_to_ride %>%
              dplyr::select(file, short_name, land),
            by = join_by(file == file)) %>%
  left_join(entities_extra %>%
              dplyr::select(short_name, duration, category_code,
                            scope_and_scale_code),
            by = join_by(short_name == short_name)) %>%
  mutate(start_wait = datetime - as.difftime(SACTMIN, units = "mins"),
         hour = hour(start_wait),
         mins = minute(start_wait),
         tod = hour + mins/60,
         inSession = as.numeric(str_remove(inSession, "%")),
         DAYOFWEEK = as.factor(DAYOFWEEK),
         YEAR = as.factor(YEAR),
         WDW_TICKET_SEASON = as.factor(WDW_TICKET_SEASON),
         scope_and_scale_code = as.factor(scope_and_scale_code)
         ) %>%
  #dplyr::select(-hour, -mins) %>%
  mutate(CapacityLostWGT_AK = scale(CapacityLostWGT_AK, center = TRUE,
                                    scale = TRUE),
         CapacityLostWGT_MK = scale(CapacityLostWGT_MK, center = TRUE,
                                    scale = TRUE),
         CapacityLostWGT_EP = scale(CapacityLostWGT_EP, center = TRUE,
                                    scale = TRUE),
         CapacityLostWGT_HS = scale(CapacityLostWGT_HS, center = TRUE,
                                    scale = TRUE),
         CapacityLostWGT = (CapacityLostWGT_AK + CapacityLostWGT_MK +
                              CapacityLostWGT_EP + CapacityLostWGT_HS)/4,
         tod2 = tod * tod, #time of day squared
         land = as.factor(land),
         category_code = as.factor(category_code),
         short_name = as.factor(short_name),
         weekend = 1*(DAYOFWEEK %in% c("6", "7")))
covariates <- data.frame(
  covariate = c("weekend", "HOLIDAYPX", "WDW_TICKET_SEASON",
                "WDWMEANTEMP", "inSession", "MKPRDDAY", "CapacityLostWGT",
                "duration", "category_code", "scope_and_scale_code", "tod"),
  label = c("weekend", "holidayPX", "ticket_season",
            "mean_temp", "schools", "nb_parades", "capacity_lost",
            "duration", "category_code", "scope_and_scale_code", "tod"),
  expl = c("Whether it is a weekend date (TRUE) or not (FALSE). Day of the week was also
           tested in the model exploration phase.",
           "Proximity to Holiday (2-directional) (in days)",
           "Walt Disney World Single Day Price Type (peak, regular, or value)",
           "Average temperature this day",
           "The percentage of schools in session",
           "Number of Daytime Parades at Magic Kingdom",
           "Total hourly capacity lost on that park day, weighted by attraction popularity. This covariate is centered and normalized.",
           "Duration of attraction in minutes",
           "The type of attraction (character_greeting, continuous_show, or ride)",
           "The type of attraction (headliner, major_attraction, minor_attraction, or  super_headliner)",
           "Time of day (in hours)"),
  type = c("binary", "numeric", "discrete", "numeric", "numeric", "numeric",
           "numeric", "numeric", "discrete", "discrete", "numeric"))
covariates %>%
  dplyr::select(label, type) %>%
  kable(col.names = c("Covariate", "Type"),
        format = "latex") %>%
  kableExtra::kable_styling(font_size = 9)
```

## Model Results

```{r eval = TRUE}
waiting_times_reg <- waiting_times_reg[complete.cases(waiting_times_reg),]
mod <- readRDS(file = find_root_file("data", "model.rds",
                                     criterion =
                                       has_file(
                                         "stat_cons_disney.Rproj")))
# Set MCMC parameters
nchains <- 3 # number of chains
niter <- 2000 # number of iterations (incl. burn-in, see next)
burnin <- niter / 4 # number of initial samples to remove (= burn-in)
nparallel <- nchains # number of cores for parallel computing
thinning <- 1 #
s <- summary(mod, prob = 0.90)
#pred <- predict(mod)
#save(pred, file = find_root_file("data", "prediction.Rdata",
                                     # criterion =
                                     #   has_file(
                                     #     "stat_cons_disney.Rproj")))
load(file = find_root_file("data", "prediction.Rdata",
                                     criterion =
                                       has_file(
                                         "stat_cons_disney.Rproj")))
```

```{r fixed, fig.cap = "Results for the fixed effects model. The errorbars show 30%, 60%, and 90% confidence intervals.", message = FALSE, warning = FALSE}
s90 <- summary(mod, prob = 0.90)
s60 <- summary(mod, prob = 0.60)
s30 <- summary(mod, prob = 0.30)


lbl <- rownames(s90$fixed)
for (x in seq_len(nrow(covariates))) {
  lbl <- str_replace(lbl,
                     covariates[x, "covariate"],
                     covariates[x, "label"])
}
labels <- 
  data.frame(raw = rownames(s90$fixed),
             label = lbl)

p <- s90$fixed %>% 
  left_join(s30$fixed) %>%
  left_join(s60$fixed) %>%
  cbind(data_frame(par = rownames(s90$fixed))) %>%
  mutate(color = ifelse(`l-90% CI` > 0, "red",
                       ifelse(`u-90% CI` < 0, "blue",
                              "black"))) %>%
  ggplot(aes(x = par, y = Estimate, color = color)) +
  geom_point(size = 2) + 
  geom_linerange(aes(ymin = `l-90% CI`, ymax = `u-90% CI`), linewidth = 0.5) +
  geom_linerange(aes(ymin = `l-60% CI`, ymax = `u-60% CI`), linewidth = 1) +
  geom_linerange(aes(ymin = `l-30% CI`, ymax = `u-30% CI`), linewidth = 1.5) +
  geom_hline(aes(yintercept = 0), color = "black") +
  coord_flip() +
  theme_bw() +
  scale_color_manual("",
                       breaks = c("blue", "black", "red"),
                       values = c("#005AB5", "black", "#DC3220")) +#colorblind friendly
  theme(legend.position = "none") +
  scale_x_discrete(name = "",
                   limits = labels$raw,
                   labels = labels$label)
p
```

##

```{r random, fig.cap = "Random effect for each of the attractions, sorted from highest to lowest difference between posted and actual waiting time."}
sd_mean <- mod %>%
  spread_draws(sd_short_name__Intercept, ndraws = 1000, seed = 123) %>%
  summarise(mean_sd = mean(sd_short_name__Intercept)) %>%
  pull()

# Neem random effects en plot
mod %>%
  spread_draws(r_short_name[short_name,], ndraws = 1000, seed = 123) %>%
  ungroup() %>%
  mutate(short_name = reorder(short_name, r_short_name)) %>%
  ggplot(aes(x = r_short_name, y = short_name)) +
    geom_vline(xintercept = 0, color = "darkgrey", linewidth = 1) +
    geom_vline(xintercept = c(sd_mean * qnorm(0.05), sd_mean * qnorm(0.95)),
               color = "darkgrey", linetype = 2) +
    stat_halfeye(point_interval = "median_qi", .width = 0.9, size = 2/3,
                 fill = "cornflowerblue") +
  theme_bw() +
  ylab("Attraction") +
  xlab("Random intercept")
```

# Conclusion

## Discussion

- Posted waiting times are, on average quite accurate.
- More likely that the posted waiting time overestimates the actual waiting time
- Improve predictions during weekends and in the evening, for headliners and character greetings

## Recommendations for future research

- Add data on throughput per attraction and total number of visitors.
- Reduce missingness in the data
- Variable selection techniques
