---
documentclass: article
fontsize: 10pt
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_document2: 
    fig_caption: true
    toc: false
    latex_engine: xelatex
    includes:
      in_header: preamble.sty
      before_body: titlepage.sty
bibliography: references.bib  
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE, appendix=TRUE}
#---------------------------load all packages----------------------------------#
library(tidyverse)
library(rprojroot)
library(kableExtra)
library(git2rdata)
library(patchwork)

knitr::opts_chunk$set(echo = FALSE, message = FALSE)
knitr::opts_knit$set(
  root.dir = find_root(criterion = has_file("stat_cons_disney.Rproj")))

#----------------------------Load all necessary data---------------------------#
waiting_times <- read_vc(file = "waiting_times",
                         root = find_root_file("data",
                                               criterion =
                                                 has_file(
                                                   "stat_cons_disney.Rproj"))) #SPOSTMIN:  Value -999 indicates the attraction was closed
file_to_ride <- read_vc(file = "file_to_ride",
                        root = find_root_file("data",
                                              criterion =
                                                has_file(
                                                  "stat_cons_disney.Rproj")))
```

# Introduction

This report investigates the accuracy of the posted waiting times for different Disneyworld attractions. For transparency, all source code for this report and the data analysis can be found a github repository [@Carmen_Disney_world_waiting].
More concrete, we will look at the difference between the posted waiting time when a customer start queueing and the actually realised waiting time.
Although the waiting times for `r n_distinct(file_to_ride$file)` attractions are available, we will focus on `r n_distinct(file_to_ride %>% dplyr::filter(!is.na(code)) %>% dplyr::pull(file))` attractions for which metadata is available.
Some attractions have more than one opening date.
Since it is unclear whether this means that there is more than one attractions with the same name or whether it pertains to a renewal of the attraction, we assume the last opening date is the only relevant one.

```{r}
#filter the metadata to only keep attractions with metadata and only one row per file (keep last opening date):
file_to_ride <- file_to_ride %>%
  dplyr::filter(!is.na(code)) %>%
  arrange(file, opened_on) %>%
  group_by(file, short_name) %>%
  slice_tail() %>%#keep only the last occurance
  ungroup()
```

To work with the waiting times, some assumptions were made:

1. The posted times are supposed to inform customers on the expected time that they will queue if they *start* queueing at any given time. This means that the prediction model that Disney world uses, tries to predict the time that a customer will have to wait **at the moment he starts queueing**.
2. The posted times do not change between two registered posted times in the data.
3. The actual waiting time of a customer is subtracted from the time when his waiting time was registered (assuming this is the time when he entered the ride) to know the time he started waiting. We then compare the posted waiting time at this calculated time with his actual waiting time.
4. Very little actual waiting times are registered each day for each attraction \@ref(fig:nb-waiting-times-reg). We assume that the customers whose actual waiting time is registered are randomly selected.

# Data exploration

In total, we have `r nrow(waiting_times)` actual waiting times over all `r n_distinct(file_to_ride$file)` attractions, registered between `r min(waiting_times$date)` and `r max(waiting_times$date)`.
Removing waiting times for which the posted waiting time was either missing or "-999" (meaning the ride was technically closed), `r sum(!is.na(waiting_times$posted_at_start_wait) &  waiting_times$posted_at_start_wait != -999)` waiting times remain.
Additionally, there are some outliers in the realized waiting times.
Removing all negative realized waiting times and waiting times larger than 360 minutes, `r sum(!is.na(waiting_times$posted_at_start_wait) &  waiting_times$posted_at_start_wait != -999 & waiting_times$SACTMIN >= 0 & waiting_times$SACTMIN <= 360)` waiting times remain.
It should be noted that we don't have the same amount of waiting times for each attractions due to the fact that the time period of data collection \@ref(fig:date-collection) and the number of collected waiting times per day \@ref(fig:nb-waiting-times-reg) differs per attraction.

```{r}
#keep only waiting times with corresponding posted waiting time:
waiting_times <- waiting_times %>%
  dplyr::filter(!is.na(posted_at_start_wait) &  posted_at_start_wait != -999 &
                  SACTMIN >= 0 & SACTMIN <= 360) %>%
  mutate(diff = posted_at_start_wait - SACTMIN)
set.seed(2024)
sample <- waiting_times %>%
  slice_sample(n = 10000) 
```


We explore the difference between the actual and posted waiting times in Figure \@ref(fig:waiting-exploration).
In the top left figure, the posted waiting time overestimated the actual waiting time for waiting times above the diagonal line.
The color corresponds to the difference between the posted and actual waiting time.
The top right figure also clearly shows that it is more likely that the actual waiting time is overestimated than underestimated (it is overestimated in approximately `r round(100*sum(sample$diff>0) / nrow(sample))`% of cases)).
It is likely that this tendency to overestimate the waiting time is a tactical choice since it has been proven that people generally get more agitated if they need to wait longer than expected.
Since the proposed waiting time will strongly influence the expected waiting times, it makes sense to overestimate rather than to underestimate the waiting time (see for instance @maister1984psychology and @chu2019psychology).
Th bottom figures show the mean and 95% confidence interval of the actual waiting time (left) and the difference between the proposed and actual waiting time (right) over the day, depending on when the client started to wait.
In the appendix, each of these figures are shown for each of the attractions separately (Figure \@ref(fig:waiting-exploration-app1) to \@ref(fig:waiting-exploration-app4)).

```{r waiting-exploration, fig.cap = "Exploration of actual and posted waiting time and the difference between both."}
#keep only waiting times with corresponding posted waiting time:
p1 <- sample %>%
  ggplot() +
  geom_line(data = data.frame(x = c(0, max(sample$SACTMIN)),
                              y = c(0,max(sample$posted_at_start_wait))),
            aes(x = x, y = y)) +
  geom_point(aes(x = SACTMIN, y = posted_at_start_wait, color = diff),
             alpha = 0.5) +
  scale_color_gradientn("Difference",
                        colours =
                          c("yellow", "orange", "red", "darkred")) +
  xlab("Actual waiting time") +
  ylab("Posted waiting time") +
  guides(colour = "none") +
  theme_bw()
bin_width <- 10
nbins = ceiling((max(sample$diff) - min(sample$diff))/bin_width)
colfunc <- colorRampPalette(c("yellow", "orange", "red", "darkred"))

p2 <- sample %>%
  ggplot() +
  geom_histogram(aes(x = diff, y = after_stat(count / sum(count))),
                 binwidth = bin_width,
                 fill = colfunc(nbins),
                 alpha = 0.8) +
  geom_vline(aes(xintercept = 0), color = "black") +
  scale_y_continuous(labels = scales::percent) +
  ylab("Percentage of waiting times") +
  xlab("Posted - actual waiting time") +
  theme_bw()

p4 <- waiting_times %>%
  mutate(
    start_wait = datetime - as.difftime(SACTMIN, units = "mins"),
    hour = hour(start_wait)) %>%
  group_by(hour) %>%
  summarize(mean = mean(diff),
            st_dev = sd(diff),
            nb = n(),
            lower = mean - 1.95 * st_dev/sqrt(nb),
            upper = mean + 1.95 * st_dev/sqrt(nb)) %>%
  ungroup() %>%
  filter(hour >= 6 & hour <= 23) %>%
  ggplot(aes(x = hour, y = mean)) +
  geom_ribbon(aes(ymin = lower, ymax = upper),
              alpha = 0.5) +
  geom_line() +
  theme_bw() +
  scale_x_continuous(breaks = seq(6, 23, 3)) +
  ylab("Proposed - actual waiting time (min)") +
  xlab("Time of day")
p3 <- waiting_times %>%
  mutate(
    start_wait = datetime - as.difftime(SACTMIN, units = "mins"),
    hour = hour(start_wait)) %>%
  group_by(hour) %>%
  summarize(mean = mean(SACTMIN),
            st_dev = sd(SACTMIN),
            nb = n(),
            lower = mean - 1.95 * st_dev/sqrt(nb),
            upper = mean + 1.95 * st_dev/sqrt(nb)) %>%
  ungroup() %>%
  filter(hour >= 6 & hour <= 23) %>%
  ggplot(aes(x = hour, y = mean)) +
  geom_ribbon(aes(ymin = lower, ymax = upper),
              alpha = 0.5) +
  geom_line() +
  theme_bw() +
  scale_x_continuous(breaks = seq(6, 23, 3)) +
  ylab("Actual waiting time (min)") +
  xlab("Time of day")


p1 + p2 + p3 + p4
```



\clearpage

\appendix

# Appendix

## Number of registered waiting times

```{r nb-waiting-times-reg, fig.cap="Mean number of registered actual waiting times per day (and 95% conf. int.) for the different attractions.", fig.height = 4, fig.width = 8}
ndays <- waiting_times %>%
  group_by(file, date) %>%
  summarize(n = n()) %>%
  summarize(mean = mean(n),
            st_dev = sd(n),
            ndays = n()) %>%
  mutate(lower = mean - 1.95 * st_dev/sqrt(ndays),
         upper = mean + 1.95 * st_dev/sqrt(ndays)) %>%
  ungroup() %>%
  arrange(mean) %>%
  left_join(file_to_ride)
ndays %>%
  ggplot(aes(x = file, y = mean, ymin = lower, ymax = upper,
             color = land)) +
  geom_pointrange(fatten = 2) +
  scale_x_discrete(name = "Attraction", breaks = ndays$file, limits = ndays$file,
                   labels = str_wrap(ndays$short_name, width = 15)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 8),
        legend.text = element_text(size = 8),
        legend.position = "bottom",
        legend.margin = margin(0,0,0,0),
        legend.box.margin = margin(-10,-10,-10,-10)) +
  ylab("Nb registered actual waiting times per day")
```

```{r date-collection, fig.cap="Period of data collection for each attraction. Color shows the total number of registered waiting times", fig.height = 4, fig.width = 8}
start_end <- waiting_times %>%
  mutate(date = ymd(as.Date(datetime))) %>%
  group_by(file) %>%
  summarize(ymin = min(date),
            ymax = max(date),
            nb = n()) %>%
  ungroup() %>%
  mutate(ymin = as.Date(ymin),
         ymax = as.Date(ymax)) %>%
   arrange(nb) %>%
  left_join(file_to_ride)
start_end %>%
  ggplot(aes(x = file, ymin = ymin, ymax = ymax, color = nb)) +
  geom_linerange(linewidth = 2)  +
  scale_y_date(date_labels = "%m/%Y") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 8),
        legend.text = element_text(size = 8),
        legend.position = "bottom",
        legend.margin = margin(0,0,0,0),
        legend.box.margin = margin(-10,-10,0,-10)) +
  scale_x_discrete(name = "Attraction", breaks = start_end$file,
                   limits = start_end$file,
                   labels = str_wrap(ndays$short_name, width = 15)) +
  scale_color_gradientn("Nb waiting\n times",
                        colours =
                          c("yellow", "orange", "darkred"))

```

## Difference between the actual and posted waiting times per attraction

Figure \@ref(fig:waiting-exploration-app1) shows the posted versus actual waiting times. *Enchanted Tiki Run* and *Flight of Passage* stand out for the shortest and longest posted and actual waiting times. The most extremely underestimated waiting times (in yellow) can be found in the *Soarin'* and *Flight of Passage* attractions.

Figure \@ref(fig:waiting-exploration-app2) shows a histogram of the difference between the posted and actual waiting time. The percentage in the top right corner indicates the percentage of observations where the posted waiting time exceeds the actual waiting time. *Regal Carrousel* stands out since it is the only attraction where less than half of the waiting times are overextimated. 


Figure \@ref(fig:waiting-exploration-app3) shows the actual waiting time over the day. We only show the mean and confidence interval if there are at least 5 customers who started waiting at that hour of the day and whose waiting time was registered. *Flight of Passage* and *7 Dwarfs Train* stand out since they have the highest waiting times. *Flight of Passage* has consistently high waiting times while most other attractions show a daily pattern with the highest waiting times between 10AM and 3PM.

Figure \@ref(fig:waiting-exploration-app4) shows the difference between the posted and actual waiting times over the day. We only show the mean and confidence interval if there are at least 5 customers who started waiting at that hour of the day and whose waiting time was registered. Most attractions show higher overestimations of the waiting time as the day progresses (especially noticeable in *Flight of Passage*, *7 Dwarfs Train*, and *Rock Coaster*)

```{r waiting-exploration-app1, fig.cap = "Exploration of actual and posted waiting time.", fig.width = 8, fig.height = 12}
samplesize <- 2000 #keep at most this many waiting times per attraction to keep the figure manageable.
counts <- waiting_times %>%
  group_by(file) %>%
  tally()
waiting_times %>%
  dplyr::filter(file %in% (waiting_times %>%
                             group_by(file) %>%
                             tally() %>%
                             dplyr::filter(n > samplesize) %>%
                             dplyr::pull(file))) %>%
  group_by(file) %>%
  slice_sample(n = 2000) %>%
  ungroup() %>%
  rbind(waiting_times %>%
          dplyr::filter(file %in% (waiting_times %>%
                             group_by(file) %>%
                             tally() %>%
                             dplyr::filter(n <= samplesize) %>%
                             dplyr::pull(file)))) %>%
  left_join(file_to_ride) %>%
  ggplot() +
  geom_line(data = data.frame(x = c(0, max(sample$SACTMIN)),
                              y = c(0,max(sample$posted_at_start_wait))),
            aes(x = x, y = y)) +
  geom_point(aes(x = SACTMIN, y = posted_at_start_wait, color = diff),
             alpha = 0.5) +
  scale_color_gradientn("Difference",
                        colours =
                          c("yellow", "orange", "red", "darkred")) +
  facet_wrap(vars(short_name), nrow = 6) +
  xlab("Actual waiting time") +
  ylab("Posted waiting time") +
  guides(colour = "none") +
  theme_bw()
```

```{r waiting-exploration-app2, warning = FALSE, fig.cap = "Histogram of the difference between posted and actual waiting time.", fig.width = 8, fig.height = 12}
percentage_overestimated <- waiting_times %>%
  mutate(diff = posted_at_start_wait - SACTMIN) %>%
  group_by(file) %>%
  summarize(n_over = sum(diff > 0),
            perc = paste0(round(100 * n_over/n()), "%")) %>%
  left_join(file_to_ride)

waiting_times %>%
  mutate(diff = posted_at_start_wait - SACTMIN) %>%
  left_join(file_to_ride) %>%
  ggplot() +
  geom_histogram(aes(x = diff, y = after_stat(density) * bin_width),
                 binwidth = bin_width,
                 alpha = 0.8) +
  geom_vline(aes(xintercept = 0), color = "black") +
  geom_text(data = percentage_overestimated,
             aes(x = 75, y = 0.6, label = perc)) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(vars(short_name), nrow = 6) +
  ylab("Percentage of waiting times") +
  xlab("Posted - actual waiting time") +
  xlim(c(-100, 100)) +
  theme_bw()
```

```{r waiting-exploration-app3, fig.cap = "Exploration of the actual waiting timeover the day.", fig.width = 8, fig.height = 12}
waiting_times %>%
  mutate(
    start_wait = datetime - as.difftime(SACTMIN, units = "mins"),
    hour = hour(start_wait)) %>%
  group_by(hour, file) %>%
  summarize(mean = mean(SACTMIN),
            st_dev = sd(SACTMIN),
            nb = n(),
            lower = mean - 1.95 * st_dev/sqrt(nb),
            upper = mean + 1.95 * st_dev/sqrt(nb)) %>%
  ungroup() %>%
  left_join(file_to_ride) %>%
  filter(hour >= 6 & hour <= 23 & nb > 5) %>%
  ggplot(aes(x = hour, y = mean)) +
  geom_ribbon(aes(ymin = lower, ymax = upper),
              alpha = 0.5) +
  geom_line() +
  theme_bw() +
  facet_wrap(facets = vars(short_name),
             nrow = 6) +
  scale_x_continuous(breaks = seq(6, 23, 3)) +
  ylab("Actual waiting time (min)") +
  xlab("Time of day")
```

```{r waiting-exploration-app4, fig.cap = "Difference between the posted and actual waiting time.", fig.width = 8, fig.height = 12}
waiting_times %>%
  mutate(
    start_wait = datetime - as.difftime(SACTMIN, units = "mins"),
    hour = hour(start_wait)) %>%
  group_by(hour, file) %>%
  summarize(mean = mean(diff),
            st_dev = sd(diff),
            nb = n(),
            lower = mean - 1.95 * st_dev/sqrt(nb),
            upper = mean + 1.95 * st_dev/sqrt(nb)) %>%
  ungroup() %>%
  filter((hour >= 6 & hour <= 23) & nb > 5) %>% # we need at least 5 waiting times to reliably calculate conf int
  left_join(file_to_ride) %>%
  ggplot(aes(x = hour, y = mean)) +
  geom_ribbon(aes(ymin = lower, ymax = upper),
              alpha = 0.5) +
  geom_line() +
  theme_bw() +
  facet_wrap(facets = vars(short_name),
             nrow = 6) +
  scale_x_continuous(breaks = seq(6, 24, 6)) +
  ylab("Posted - actual waiting time (min)") +
  xlab("Time of day")  +
  scale_y_continuous(limits = c(-10, NA))
```

# Bibliography
